### 2021-06-11 msAggr_AnalysisCode/msAggr_seurat-v1.Rmd
Initiate R project repository

---
title: "msAggr_seurat-v1"
output: html_notebook
---

Creating new pipeline using seurat v4.0.2 available 2021.06.08

Load libraries required for Seuratv4

```{r setup}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/")
library(dplyr)
library(Seurat)
library(patchwork)
```
store session info
```{r}
sink("msAggr_seurat-v1.20210608")
sessionInfo()
sink()
```

# Define functions

```{r}
read_10XGenomics_data<-function(sample.list){
	data <- sapply(sample.list, FUN = function(x) {x = dirname(list.files(path = x, pattern = 'barcodes.tsv', full.names = TRUE, recursive = TRUE, no.. = TRUE))})
	names(data) <- sample.list
	return(data)
}
```
# A note about using SCTransform versus `ScaleData`
https://bioconductor.org/packages/3.10/workflows/vignettes/simpleSingleCell/inst/doc/batch.html#62_for_gene-based_analyses
>You can also normalize and scale data for the RNA assay. There are numerous resources on this, but Aaron Lun describes why the original log-normalized values should be used for DE and visualizations of expression quite well here:
>
>For gene-based procedures like differential expression (DE) analyses or gene network construction, it is desirable to use the original log-expression values or counts. The corrected values are only used to obtain cell-level results such as clusters or trajectories. Batch effects are handled explicitly using blocking terms or via a meta-analysis across batches. We do not use the corrected values directly in gene-based analyses, for various reasons:
>
>It is usually inappropriate to perform DE analyses on batch-corrected values, due to the failure to model the uncertainty of the correction. This usually results in loss of type I error control, i.e., more false positives than expected.
>
>The correction does not preserve the mean-variance relationship. Applications of common DE methods like edgeR or limma are unlikely to be valid.
>
>Batch correction may (correctly) remove biological differences between batches in the course of mapping all cells onto a common coordinate system. Returning to the uncorrected expression values provides an opportunity for detecting such differences if they are of interest. Conversely, if the batch correction made a mistake, the use of the uncorrected expression values provides an important sanity check.
>
>In addition, the normalized values in SCT and integrated assays don't necessary correspond to per-gene expression values anyway, rather containing residuals (in the case of the scale.data slot for each).

Mess with how to load 4 cell populations into single seurat object

SET SEED?????!!!!!

```{r}
data_files.list<-read_10XGenomics_data(data)
seurat.object.data<-Read10X(data_files.list)
seurat.object<- CreateSeuratObject(raw.data = seurat.object.data, min.cells = 3, min.genes = 200, project = projectName)
seurat.object<- create_percentMito_column(seurat.object)

```

```{r}
seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r}
plot1 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident")
plot2 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident")
plot1 + plot2
```

remove low quality cells
require: nFeature_RNA between 200 and 2500 (inclusive)
--require: percent.mt <5--???
```{r}
print(paste("original object", seurat.object, sep = " "))
seurat.object <- subset(seurat.object, 
												subset = nFeature_RNA >=200 & 
													nFeature_RNA <= 2500
												)
print(paste("new object", seurat.object, sep = " "))
```

```{r}
seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
```

Find variable features

```{r}
seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 2000)
```

`selection.method	
How to choose top variable features. Choose one of :

vst: First, fits a line to the relationship of log(variance) and log(mean) using local polynomial regression (loess). Then standardizes the feature values using the observed mean and expected variance (given by the fitted line). Feature variance is then calculated on the standardized values after clipping to a maximum (see clip.max parameter).

mean.var.plot (mvp): First, uses a function to calculate average expression (mean.function) and dispersion (dispersion.function) for each feature. Next, divides features into num.bin (deafult 20) bins based on their average expression, and calculates z-scores for dispersion within each bin. The purpose of this is to identify variable features while controlling for the strong relationship between variability and average expression.

dispersion (disp): selects the genes with the highest dispersion values`