### 2021-06-16 msAggr_AnalysisCode/msAggr_sctransform-v1.Rmd
start work on SCTransform method

---
title: "msAggr_sctransform-v1"
output: html_notebook
---

Creating new pipeline using seurat v4.0.2 available 2021.06.08

Load libraries required for Seuratv4

```{r setup}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/")
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(clustree)
```
store session info
```{r}
sink("msAggr_sct-v1.20210615")
sessionInfo()
sink()
```

# A note about using SCTransform versus `ScaleData`
https://bioconductor.org/packages/3.10/workflows/vignettes/simpleSingleCell/inst/doc/batch.html#62_for_gene-based_analyses
>You can also normalize and scale data for the RNA assay. There are numerous resources on this, but Aaron Lun describes why the original log-normalized values should be used for DE and visualizations of expression quite well here:
>
>For gene-based procedures like differential expression (DE) analyses or gene network construction, it is desirable to use the original log-expression values or counts. The corrected values are only used to obtain cell-level results such as clusters or trajectories. Batch effects are handled explicitly using blocking terms or via a meta-analysis across batches. We do not use the corrected values directly in gene-based analyses, for various reasons:
>
>It is usually inappropriate to perform DE analyses on batch-corrected values, due to the failure to model the uncertainty of the correction. This usually results in loss of type I error control, i.e., more false positives than expected.
>
>The correction does not preserve the mean-variance relationship. Applications of common DE methods like edgeR or limma are unlikely to be valid.
>
>Batch correction may (correctly) remove biological differences between batches in the course of mapping all cells onto a common coordinate system. Returning to the uncorrected expression values provides an opportunity for detecting such differences if they are of interest. Conversely, if the batch correction made a mistake, the use of the uncorrected expression values provides an important sanity check.
>
>In addition, the normalized values in SCT and integrated assays don't necessary correspond to per-gene expression values anyway, rather containing residuals (in the case of the scale.data slot for each).

Load 4 populations into single seurat object

SET SEED?????!!!!!

# Set global variables
```{r}
projectName <- "msAggr_sct"
jackstraw.dim <- 40
```

```{r}
source("msAggr_AnalysisCode/read_10XGenomics_data.R")

```

```{r}
setwd("../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = c("LSKm2", "CMPm2", "MEPm", "GMPm"))
object.data <-Read10X(data_file.list)
```

```{r}
seurat.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = projectName)
```

Clean up to free memory

```{r}
remove(seurat.object.data)
```

Add mitochondrial metadata and plot some basic features
```{r}
seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident', )
```

```{r fig.width=4, fig.height=2}
plot1 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident", pt.size = 0.01)
plot2 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident", pt.size = 0.01)
plot1 + plot2
```
remove low quality cells **
require: nFeature_RNA between 200 and 4000 (inclusive)
--require: percent.mt <5--???

\** In previous run it's possible that we'll need to limit percent.mt because of MEPm weirdness. Will have to see if the additional 2000 MEPm cells found by cellranger v6

```{r}
print(paste("original object:", nrow(seurat.object@meta.data), "cells", sep = " "))
seurat.object <- subset(seurat.object, 
												subset = nFeature_RNA >=200 & 
													nFeature_RNA <= 4000
												)
print(paste("new object:", nrow(seurat.object@meta.data), "cells", sep = " "))
seurat.object <- PercentageFeatureSet(seurat.object, pattern = '^mt-', col.name = "percent.mt")
```

## SCTransform

As previously mentioned, `SCTransform()` is best for batch correction, whereas `NormalizeData()`, `ScaleData()`, and `FindVariableFeatures()` are best for DGE. I'm struggling to make stable clusters, so I'm jumping ahead to `SCTransform()`.  
It does appear that `SCTransform()` data are stored in the SCT assay slot, which does not overwrite other normalization slots. It will become the default slot once populated.

```{r}
seurat.object <- SCTransform(seurat.object, 
														 method = "glmGamPoi", 
														 vars.to.regress = c("nCount_RNA", "nFeature_RNA")
														 ) # method = "glmGamPoi" speeds process
```

Tada! Now lets do some plotting and clustering!

linear dimensional reduction. Default are based on VariableFeatures, but can be changed

```{r}
seurat.object <- RunPCA(seurat.object, features = VariableFeatures(object = seurat.object), verbose = FALSE)
```

```{r fig.width=4, fig.height=4}
VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 2)
```

DimPlot colored by orig.ident
```{r}
DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
```

```{r}
source("msAggr_AnalysisCode/PercentVariance.R")
```

```{r, figures-side, fig.show='hold', out.width="50%"}
ElbowPlot(seurat.object, ndims = 50)
percent.variance(seurat.object@reductions$pca@stdev)
```
# Variance described by dimensionality
Number of PCs describing X% of variance

```{r}
tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))

```

# Dim33

Cover 95% variability

```{r}
ndim <- 33
```

```{r}
seurat.object <- RunUMAP(seurat.object, dims = 1:ndim)
seurat.object <- FindNeighbors(seurat.object, dims = 1:ndim)
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	seurat.object <- FindClusters(seurat.object, resolution = x)
}
```
Plotting...

```{r}
for (meta.col in colnames(seurat.object@meta.data)){
	if(grepl(pattern = ("SCT_snn_res"), x = meta.col)==TRUE){
		myplot <- DimPlot(seurat.object, 
											group.by = meta.col,
											reduction = "umap", 
											cols = colorRamps::primary.colors(n = length(levels(seurat.object@meta.data[[meta.col]])))
											) + 
			ggtitle(paste0("msAggr sct dim31 res", gsub("SCT_snn_res", "", meta.col) ))
		plot(myplot)
	}
}
```

```{r}
saveRDS(seurat.object, file = "msAggr_sctdim33.rds")
```

## Evaluate cluster stability
Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.

Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.

### Clustree
Previously my favourite has been Clustree, which gives a nice visual
NB: For some reason `clustree::clustree()` didn't work, whereas `library(clustree)` followed by `clustree()` did.

```{r fig.height=5}
clustree(seurat.object, prefix = "SCT_snn_res.", node_colour = "sc3_stability") + 
	scale_color_continuous(low = 'red3', high = 'white')
```

Basically no stability difference compared to before.

# Dim22

Cover 90% variability

```{r}
ndim <- 22
```

```{r}
seurat.object <- RunUMAP(seurat.object, dims = 1:ndim)
seurat.object <- FindNeighbors(seurat.object, dims = 1:ndim)
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	seurat.object <- FindClusters(seurat.object, resolution = x)
}
```
Plotting...

```{r}
for (meta.col in colnames(seurat.object@meta.data)){
	if(grepl(pattern = ("SCT_snn_res"), x = meta.col)==TRUE){
		myplot <- DimPlot(seurat.object, 
											group.by = meta.col,
											reduction = "umap", 
											cols = colorRamps::primary.colors(n = length(levels(seurat.object@meta.data[[meta.col]])))
											) + 
			ggtitle(paste0("msAggr sct dim", ndim, "res", gsub("SCT_snn_res", "", meta.col) ))
		plot(myplot)
	}
}
```

```{r}
saveRDS(seurat.object, file = paste0("msAggr_sctdim", ndim, ".rds"))
```

## Evaluate cluster stability
Must ensure we have the right cluster stability, that is, cells that start in the same cluster tend to stay in the same cluster. If your data is over-clustered, cells will bounce between groups.

Following [this tutorial by Matt O.].https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92.

### Clustree
Previously my favourite has been Clustree, which gives a nice visual
NB: For some reason `clustree::clustree()` didn't work, whereas `library(clustree)` followed by `clustree()` did.

```{r fig.height=5}
clustree(seurat.object, prefix = "SCT_snn_res.", node_colour = "sc3_stability") + 
	scale_color_continuous(low = 'red3', high = 'white')
```

Hard to tell if dim33 or dim22 is better, but if *feels* like dim22 has less cross-over between res0.5 and res1.0. Will continue at dim22, and see how things change between res0.5 and res1.0

```{r}
# Number of filtered cells left in each pop
sapply(c("LSKm2", "CMPm2", "MEPm", "GMPm"), function(x) (c(nrow(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x,]))))
```
## MitoC in

```{r}
for (x in c("LSKm2", "CMPm2", "MEPm", "GMPm")){
	h = hist(seurat.object@meta.data[seurat.object@meta.data$orig.ident == x, 'percent.mt'], breaks = 30, plot = FALSE)
	h$density = h$counts/sum(h$counts)*100
	plot(h,freq=FALSE, main =  paste(x, "percent mitoC"), xlab = "percent mitoC", ylab = "Frequency")
}