### 2021-06-22 CMPSubset.Rmd
Added CMP subset

---
title: "CMPSubset"
output: html_notebook
---

Creating new pipeline using seurat v4.0.2 available 2021.06.08

Load libraries required for Seuratv4

```{r setup}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/")
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
# library(clustree)
```
store session info
```{r}
sink("CMP-v1.20210616.txt")
sessionInfo()
sink()
```

# Set global variables
```{r}
projectName <- "CMP"
jackstraw.dim <- 40
```

```{r}
source("msAggr_AnalysisCode/read_10XGenomics_data.R")
source("msAggr_AnalysisCode/PercentVariance.R")
source("msAggr_AnalysisCode/Mouse2Human_idconversion.R")
```

```{r}
setwd("../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
object.data <-Read10X(data_file.list)
```

```{r}
seurat.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = projectName)
```

Clean up to free memory

```{r}
remove(object.data)
```

Add mitochondrial metadata and plot some basic features
```{r}
seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident', )
```

```{r}
plot1 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident", pt.size = 0.01)
plot2 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident", pt.size = 0.01)
plot1 + plot2
```

We don't have to worry about comparing library depths, so we'll just do normalization/Scale data

remove low quality cells
require: nFeature_RNA between 200 and 4000 (inclusive)
require: percent.mt <=5

```{r}
print(paste("original object:", nrow(seurat.object@meta.data), "cells", sep = " "))
seurat.object <- subset(seurat.object, 
												subset = nFeature_RNA >=200 & 
													nFeature_RNA <= 4000 & 
													percent.mt <= 5
												)
print(paste("new object:", nrow(seurat.object@meta.data), "cells", sep = " "))
```

```{r}
seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
```

Find variable features
```{r fig.width = 5, fig.height = 2}
seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 2000)
top10 <- head(VariableFeatures(seurat.object), 10)
plot1 <- VariableFeaturePlot(seurat.object)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2

```

Scale data (linear transformation)

```{r}
all.genes <- rownames(seurat.object)
seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("nFeature_RNA", "nCount_RNA"))
```

```{r}
seurat.object <- RunPCA(seurat.object, features = VariableFeatures(seurat.object), ndims.print = 1:5, nfeatures.print = 5)
```

```{r}
DimPlot(seurat.object, reduction = "pca")
VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)

```

Calculate dimensionality
```{r, figures-side, fig.show='hold', out.width="50%"}
ElbowPlot(seurat.object, ndims = 50)
percent.variance(seurat.object@reductions$pca@stdev)
```
Number of PCs describing X% of variance

```{r}
tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))

```

# Cluster cells

Let's take 90% of variance
```{r}
ndims <- 33
```

```{r}
seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
seurat.object <- FindClusters(seurat.object, resolution = 0.5)
```
## Generate UMAP
```{r}
seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)
```
Plot UMAP

```{r}
DimPlot(seurat.object)
```

## Add cluster IDs from Seurat v1

Exported cell IDs for clusters 3, 17, 10, 11 from Seurat v1. Will add these IDs as a metadata column.  
Create column "clust.ID" and populate with 0's. Then import IDs for clusters

```{r}
clust3.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust17.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust10.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust11.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
```

Add new metadata column
```{r}
seurat.object@meta.data['clust.ID'] <- 0
head(seurat.object@meta.data)
```

now map new ids
```{r}
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust3.cells] <- 3
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust17.cells] <- 17
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust10.cells] <- 10
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust11.cells] <- 11
```

do numbers make sense?

```{r}
nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 10,])
nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 11,])
nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 17,])
nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 3,])
```

make enough sense!

## Find old cells on UMAP

time for the super scarey moment to see if the cells from seuratv1 still cluster together on in seurat v4

```{r}
DimPlot(seurat.object,
				reduction = "umap",
				group.by = "clust.ID", 
				cols = c("gray", "orange", "blue", "red", "green"))
```

Is this success?? It may be!! At the very least they're clustering togehter...  
print out the picture
```{r}
png(filename = "ClusterIDs_on_newCMP.png", height = 800, width = 800)
DimPlot(seurat.object,
				reduction = "umap",
				group.by = "clust.ID",
				cols = c("gray", "orange", "blue", "red", "green"), 
				pt.size = 1)
dev.off()
```

Let's see if we can get some gene expression profiles on these...
```{r, fig.height=10, fig.width=18}
gene.list <- c("Gata1", "Gata2", "Pf4", "Dntt", "Mpo", "Meis1", "Irf8", "Elane", "Fli1", "Zfpm1")
VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
```

Write out picture w/ & w/o dots
```{r}
png(filename = "Fav_genes_on_newCMP.png", height = 800, width = 1600)
VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0, cols = c("gray", "orange", "blue", "red", "green"))
dev.off()

png(filename = "Fav_genes_on_newCMP-dotted.png", height = 800, width = 1600)
VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
dev.off()

```

# DGE

Let's do some cluster analyses and see if we can find these populations in our new analysis.
## Ideal resolution...? (Is this a thing?)

```{r}
color.palette <- c(
	"coral",
	"chartreuse4",
	"goldenrod1",
	"cadetblue1",
	"burlywood",
	"brown",
	"brown1",
	"blue",
	"blue4",
	"azure3",
	"aquamarine",
	"antiquewhite",
	"cadetblue",
	"gold3",
	"black",
	"darkgreen",
	"deeppink",
	"darkviolet",
	"darkturquoise",
	"darkslategray",
	"darksalmon",
	"darkorchid1",
	"darkolivegreen2",
	"forestgreen",
	"dodgerblue",
	"green",
	"lightpink",
	"lightcoral",
	"khaki1",
	"maroon",
	"peru",
	"lightseagreen",
	"lightsalmon",
	"plum",
	"moccasin",
	"tan",
	"tan1"
)
```

```{r}
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	seurat.object <- FindClusters(seurat.object, resolution = x)
}
```

```{r}
for (meta.col in colnames(seurat.object@meta.data)){
	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
		myplot <- DimPlot(seurat.object, 
											group.by = meta.col,
											reduction = "umap", 
											cols = color.palette
											) + 
			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
		plot(myplot)
	}
}
```

For each resolution, what percentage of cells in each cluster are enriched for one of our clust.IDs?

```{r}
temp <- seurat.object@meta.data
df <- data.frame(matrix(ncol = 4, nrow = length(levels(temp$RNA_snn_res.0.5))))
colnames(df) <- c("c3", "c17", "c10", "c11")
rownames(df) <- sort(as.numeric(levels(temp$RNA_snn_res.0.5)))
tot.clus <- nrow(temp[temp$RNA_snn_res.0.5 == "0",])
num.3 <- nrow(temp[(temp$RNA_snn_res.0.5 == "0") & (temp$clust.ID == "3"),])
pct.3 <- as.integer(num.3/tot.clus * 100)
df["0", "c3"]<- pct.3
```

Test: what percentage of each new clusterID matches one of the older clusters?
```{r}
for (meta.col in colnames(seurat.object@meta.data)){
	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
		new.clusters <- sort(as.numeric(levels(seurat.object@meta.data[[meta.col]])))
		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
		colnames(enrich.df) <- c(3, 17, 10, 11)
		rownames(enrich.df) <- new.clusters
		meta.df <- seurat.object@meta.data
		for(row.id in rownames(enrich.df)){
			tot.clus <- nrow(meta.df[meta.df[[meta.col]] == row.id,])
			for(col.id in colnames(enrich.df)){
				num.x <- nrow(meta.df[(meta.df[[meta.col]] == row.id) & (meta.df$clust.ID == col.id),])
				pct.x <- as.integer(num.x / tot.clus *100)
				# print(pct.x)
				enrich.df[row.id, col.id] <- pct.x
			}
		}
		colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0("oldcluster", x))
		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
		xlsx::write.xlsx(enrich.df, file = "PctOfNewClustersOverlappingOldClusters.xlsx", sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
		print(enrich.df)
	}
}

```

Used the exce doc to do some fancy conditional formatting. Old cluster 17 is pretty dispersed until you it resolution 2.5. Otherise, cells in old cluster 17 do not constitute more than 40% of any cells in the new clusters.  
As far as I can see, the two approaches are to do DGEof new CMP w/ resolution = 2.5, AND/OR do DGe using older cluster IDs. Sure seems to make sense to do both...

# DGE w/ resolution = 2.5
Strt with comparing all clusters against all other clusters
Write out cluster info

calculate `FindAllMarkers()` for different idents and save to new file
```{r}
ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1", "RNA_snn_res.1.5", "RNA_snn_res.2", "RNA_snn_res.2.5", "clust.ID")
for(tested.ident in ident.list){
	Idents(seurat.object) <- tested.ident
	all.markers <- FindAllMarkers(seurat.object)
	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
									 file = paste0(projectName, "_FindALLMarkers_res2.5.xlsx"), 
									 sheetName = tested.ident, 
									 col.names = TRUE, 
									 row.names = FALSE, 
									 append = TRUE)
}
```

Create `FindAllMarkers()` lists for GSEA
```{r}
Idents(seurat.object) <- "RNA_snn_res.2.5"
res.2.5.allmarkers <- FindAllMarkers(seurat.object)
```

Map HGNC symbols
```{r}
Mouse2HumanTable <- Mouse2Human(res.2.5.allmarkers$gene)

HGNC <- with(Mouse2HumanTable, Mouse2HumanTable$HGNC[match(res.2.5.allmarkers$gene, Mouse2HumanTable$MGI)])
head(res.2.5.allmarkers)
res.2.5.allmarkers$HGNC <- HGNC
tail(res.2.5.allmarkers)
sig.res.2.5 <- res.2.5.allmarkers[res.2.5.allmarkers$p_val_adj <= 0.05, ]
sig.res.2.5 <- sig.res.2.5[c("avg_log2FC", "HGNC", "cluster")]
sig.res.2.5 <- sig.res.2.5[!(sig.res.2.5$HGNC == "" | is.na(sig.res.2.5$HGNC)),] # GSEA will fail if there are any blanks or NAs in the table
sig.res.2.5 <- sig.res.2.5[]

```

```{r}
for(cluster in unique(sig.res.2.5$cluster)){
	print(paste("writing cluster", cluster))
	new.table <- sig.res.2.5[sig.res.2.5$cluster == cluster, c("HGNC", "avg_log2FC")]
	new.table <- new.table[order(-new.table$avg_log2FC), ]
	write.table(new.table, file = paste0("RankList_res2.5_findAll_hgnc/res.2.5cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", )

}
```

calculate `FindMarkers()` that distinguish each cluster (might overlab between clusters)
```{r}
ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1", "RNA_snn_res.1.5", "RNA_snn_res.2", "RNA_snn_res.2.5", "clust.ID")
for(tested.ident in ident.list){
	for (cluster in sort(as.numeric(levels(seurat.object@meta.data[[tested.ident]])))){
	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
									 file = paste0(projectName, "_FindMarkers_", gsub("RNA_snn_", "", tested.ident), ".xlsx"), 
									 sheetName = paste0("clst", cluster), 
									 col.names = TRUE, 
									 row.names = TRUE, 
									 append = TRUE)
}
}
```

```{r}
for (cluster in sort(as.numeric(levels(seurat.object@meta.data$RNA_snn_res.2.5)))){
	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
									 file = paste0(projectName, "_FindMarkers_res2.5.xlsx"), 
									 sheetName = paste0("clst", cluster), 
									 col.names = TRUE, 
									 row.names = TRUE, 
									 append = TRUE)
}
```

## Combine clusters that might represent old cluster ids

# DGE w/ metadata against clust.ID against "0"
reset ident as "clust.ID" and rerun `FindAllMarkers()`
```{r}
	Idents(seurat.object) <- "clust.ID"
	all.markers <- FindAllMarkers(seurat.object)
	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
									 file = paste0(projectName, "_FindALLMarkers_clustID.xlsx"), 
									 sheetName = "clustID", 
									 col.names = TRUE, 
									 row.names = FALSE, 
									 append = TRUE)
```

```{r}
# Idents(seurat.object) <- "clust.ID"
for (cluster in unique(seurat.object@meta.data$clust.ID)){
	print(cluster)
	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
									 file = paste0(projectName, "_FindMarkers_clustID.xlsx"), 
									 sheetName = paste0("oldclust", cluster), 
									 col.names = TRUE, 
									 row.names = TRUE, 
									 append = TRUE)
}

```

# Cell cycle analysis
Just so we know what we're working with

```{r}
s.genes <- cc.genes.updated.2019$s.genes
g2m.genes <- cc.genes.updated.2019$g2m.genes
```