### 2021-10-14 MergeBiomarkWith10X/MergeBiomarkWith10X.Rmd
Completed script to combine 10X and biomark into tech.integrated seurat objects and export biomark flow data as ridgeplots

---
title: "MergeBiomarkWith10X"
output:
  html_document:
    df_print: paged
---

# Updates
## 20211011
Have successfully co-clustered LSK, CMP, and MEP scCells with respective biomark populations. Now have to determine if the fluorophore data supports this.  
Will load the fluorophore data in as metadata, and try to plot the fluor metadata based on the sc/biomark clusters. not sure if the NAs in sc will be a problem, but can subset to the biomark object and try that way.

## 20211007
Gonna crank through LSK and MEP, then go back to the fluorophore data

## 20211005
Compared to archive 1, this version imports Biomark Ct spreadsheets that do not include fluorophore data

# Setup
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/MergeBiomarkWith10X/")
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(clustree)
```

## Set global variables
```{r}
pop.of.interest <- "LSK"
cellranger.folder <- "m2"
fav.var <- 95
object.res <- ".0.5"
ident.list <- c("integrated_snn_res.0.5", "integrated_snn_res.1")
projectName <- paste0("MergeBiomarkWith10X_", pop.of.interest)
```
source code
```{r}
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/Mouse2Human_idconversion.R")
```
store session info
```{r }
sink(paste0(projectName, "_sessionInfo.txt"))
sessionInfo()
sink()
```

# Create seurat object

## Load biomark data
Try to read in Biomark Data

```{r}
biomark.data <- read.table("/Users/heustonef/Desktop/10XGenomicsData/BioMark/Biomark_CT.txt", sep = "\t", header = TRUE, row.names = 1)
biomark.data[is.na(biomark.data)] <- 0

# Limit to biomark pop subset
biomark.data <- biomark.data[grepl(pop.of.interest, rownames(biomark.data)) == TRUE,]
dim(biomark.data)
```
```{r}
# fix non-MGI gene names
colnames(biomark.data)[colnames(biomark.data) == 'Cd117'] <- 'Kit'
colnames(biomark.data)[colnames(biomark.data) == 'Cd123'] <- 'Il3ra'
colnames(biomark.data)[colnames(biomark.data) == 'H2.Aa'] <- 'H2-Aa'
colnames(biomark.data)[colnames(biomark.data) == 'H2.Ab1'] <- 'H2-Ab1'
biomark.data <- subset(biomark.data, select = -Kng2)
```
Convert to pseudoexpression data
```{r}
# generate list of reference Ct averages, calculating average for ref genes where ref_gene Ct > 0
ref_list <- c("Actb", "B2m", "Kit")
biomark.data["refCt"] <- NA

for(i in 1:nrow(biomark.data)){
	ref_cts <- c()
	for(j in ref_list){
		if(biomark.data[i, j] > 0){
			ref_cts <- c(ref_cts, biomark.data[i, j])
		}
	}
	biomark.data[i, "refCt"] <- mean(ref_cts)
}

# generate dCt col
dbiomark.data <- biomark.data[, 1:ncol(biomark.data) -1]
dbiomark.data[dbiomark.data == 0] <- NA
dbiomark.data <- dbiomark.data - biomark.data[, "refCt"]

# create pseudo expression matrix
pseudo.biomark <- 2^-dbiomark.data

```

```{r warning=FALSE, message=FALSE}
biomark.meta <- data.frame(stringr::str_split_fixed(rownames(pseudo.biomark), "_", 3)[,2:3], row.names = rownames(pseudo.biomark))
colnames(biomark.meta) <- c("wid", "pid")

pseudo.biomark.T <- data.table::transpose(pseudo.biomark)
colnames(pseudo.biomark.T) <- rownames(pseudo.biomark)
rownames(pseudo.biomark.T) <- colnames(pseudo.biomark)

biomark.meta['nFeature_RNA'] <- NA
for(cell in colnames(pseudo.biomark.T)){
	ngenes <- length(which(!(pseudo.biomark.T[colnames(pseudo.biomark.T > 0) == cell])))
	biomark.meta[rownames(biomark.meta) == cell, 'nFeature_RNA'] <- ngenes
}

```

```{r}
pseudo.biomark.T[is.na(pseudo.biomark.T)] <- 0
pseudo.biomark.T <- as.sparse(pseudo.biomark.T)
biomark.object <- CreateSeuratObject(pseudo.biomark.T, project = "biomark", assay = "RNA", meta.data = biomark.meta)
```

## Load SC data

```{r warning=FALSE}
setwd("~/Desktop/10XGenomicsData/cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = paste0(pop.of.interest, cellranger.folder))
object.data <-Read10X(data_file.list)
sc.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = "10X")
```

Keep only genes in both objects

```{r}
shared.features <- intersect(rownames(biomark.object@assays$RNA@data), rownames(sc.object@assays$RNA@data))
sc.object <- subset(sc.object, features = shared.features)
biomark.object <- subset(biomark.object, features = shared.features)
sc.object@meta.data["wid"] <- '10X'
sc.object@meta.data["pid"] <- "p14"

# Check that there's a clean overlap
setdiff(rownames(biomark.object@assays$RNA@data), rownames(sc.object@assays$RNA@data))
```

Clean up to free memory

```{r}
remove(object.data, biomark.data, biomark.meta)
```

## merge objects
```{r}
seurat.object <- merge(sc.object, biomark.object, project = 'mergebiomark')
seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == paste0(pop.of.interest, cellranger.folder)] <- '10X'
seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == pop.of.interest] <- 'biomark'
# seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
# seurat.object@meta.data$orig.ident <- factor(seurat.object@meta.data$orig.ident)
# seurat.object@meta.data$pid <- factor(seurat.object@meta.data$pid)
# seurat.object@meta.data$wid <- factor(seurat.object@meta.data$wid)
seurat.object
```

# Traditional normalize and scaling approaches
trying sctransform and normalizeData/scaledata
```{r}
# seurat.object <- SCTransform(seurat.object, vars.to.regress = c("wid"))
seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 10)
```

Scale data (linear transformation)

```{r}
all.genes <- rownames(seurat.object)
seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("wid", "pid", "orig.ident"))
```

### Save merged object
```{r}
# saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
```
## Run PCA
```{r}
seurat.object <- RunPCA(seurat.object, ndims.print = 1:5, nfeatures.print = 5, features = all.genes)
```

```{r}
DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
```

```{r}
VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)

```

Calculate dimensionality
```{r, figures-side, fig.show='hold', out.width="50%"}
ElbowPlot(seurat.object, ndims = 50)
percent.variance(seurat.object@reductions$pca@stdev)
```

## Calculate percent variance
```{r}
tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))

```

# scRNASeq integration	
neither sctransform nor normalizedata/scaledata gave what looked like reasonable results.

https://satijalab.org/seurat/articles/integration_introduction.html

```{r}
tech.list <- c(sc.object, biomark.object)
tech.list <- lapply(X = tech.list, FUN = function(x) {
	x <- NormalizeData(x)
	x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 10)
})
```

## identify anchors
```{r}
tech.anchors <- FindIntegrationAnchors(object.list = tech.list, anchor.features = all.genes, k.anchor = 20, k.filter = NA)
# tech.integrated <- IntegrateData(anchorset = tech.anchors, dims = 1:30, k.weight = 20, normalization.method = "SCT")
tech.integrated <- IntegrateData(anchorset = tech.anchors, k.weight = 5)
```

## tech.integrated PCA

```{r}
DefaultAssay(tech.integrated) <- "integrated"
tech.integrated <- ScaleData(tech.integrated)
tech.integrated <- RunPCA(tech.integrated, npcs = 30)
DimPlot(tech.integrated, reduction = "pca", group.by = "orig.ident", pt.size = 2)
```

so it looks like anchor-based integration is much nicer! goign to stick with this approach.

Number of PCs describing X% of variance

```{r}
tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance: ", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance: ", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance: ", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance: ", length(which(cumsum(tot.var) <= 95)))

```

# ID clusters based on different variances
## Describe 80% of variance

### Neighborhood and umap
```{r}
tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
pct.var <- 80
ndims <- length(which(cumsum(tot.var) <= pct.var))
print(ndims)

tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims)
tech.integrated <- FindClusters(tech.integrated, resolution = 0.5)
tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims)
```
find clusters

```{r}
for(x in c(0.5, 1, 1.5)){
	tech.integrated <- FindClusters(tech.integrated, resolution = x)
}
```

```{r}
for (meta.col in colnames(tech.integrated@meta.data)){
	if(grepl(pattern = "integrated_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
		myplot <- DimPlot(tech.integrated, 
											group.by = meta.col,
											reduction = "umap", 
											pt.size = 1,
											cols = color.palette) + 
			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
		plot(myplot)
		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
		plot(myplot)
		dev.off()
		myplot <- DimPlot(tech.integrated, 
											group.by = meta.col,
											reduction = "umap", 
											pt.size = 1,
											cols = color.palette) + 
			facet_wrap(meta.col) + 
			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))

		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
		plot(myplot)
		dev.off()

	}
}
```

```{r}
saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
```

### Evaluate cluster stability
```{r fig.height=5}
clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") + 
	scale_color_continuous(low = 'red3', high = 'white')
png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") +
	scale_color_continuous(low = 'red3', high = 'white')
dev.off()
```

## Describe 85% of variance

### Neighborhood and umap
```{r}
tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
pct.var <- 85
ndims <- length(which(cumsum(tot.var) <= pct.var))
print(ndims)

tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims)
tech.integrated <- FindClusters(tech.integrated, resolution = 0.5)
tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims)
```
Plot UMAP

```{r}
for(x in c(0.5, 1, 1.5)){
	tech.integrated <- FindClusters(tech.integrated, resolution = x)
}
```

```{r}
for (meta.col in colnames(tech.integrated@meta.data)){
	if(grepl(pattern = "integrated_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
		myplot <- DimPlot(tech.integrated, 
											group.by = meta.col,
											reduction = "umap", 
											pt.size = 1,
											cols = color.palette) + 
			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
		plot(myplot)
		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
		plot(myplot)
		dev.off()
		myplot <- DimPlot(tech.integrated, 
											group.by = meta.col,
											reduction = "umap", 
											pt.size = 1,
											cols = color.palette) + 
			facet_wrap(meta.col) + 
			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))

		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
		plot(myplot)
		dev.off()

	}
}
```

```{r}
saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
```

### Evaluate cluster stability
```{r fig.height=5}
clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") + 
	scale_color_continuous(low = 'red3', high = 'white')
png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") +
	scale_color_continuous(low = 'red3', high = 'white')
dev.off()
```

## Describe 90% of variance

### Neighborhood and umap
```{r}
tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
pct.var <- 90
ndims <- length(which(cumsum(tot.var) <= pct.var))
print(ndims)

tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims)
tech.integrated <- FindClusters(tech.integrated, resolution = 0.5)
tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims)
```
Plot UMAP

```{r}
for(x in c(0.5, 1, 1.5)){
	tech.integrated <- FindClusters(tech.integrated, resolution = x)
}
```

```{r}
for (meta.col in colnames(tech.integrated@meta.data)){
	if(grepl(pattern = "integrated_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
		myplot <- DimPlot(tech.integrated, 
											group.by = meta.col,
											reduction = "umap", 
											pt.size = 1,
											cols = color.palette) + 
			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
		plot(myplot)
		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
		plot(myplot)
		dev.off()
		myplot <- DimPlot(tech.integrated, 
											group.by = meta.col,
											reduction = "umap", 
											pt.size = 1,
											cols = color.palette) + 
			facet_wrap(meta.col) + 
			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))

		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
		plot(myplot)
		dev.off()

	}
}
```

```{r}
saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
```

### Evaluate cluster stability
```{r fig.height=5}
clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") + 
	scale_color_continuous(low = 'red3', high = 'white')
png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") +
	scale_color_continuous(low = 'red3', high = 'white')
dev.off()
```

## Describe 95% of variance

### Neighborhood and umap
```{r}
tot.var <- percent.variance(tech.integrated@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
pct.var <- 95
ndims <- length(which(cumsum(tot.var) <= pct.var))
print(ndims)

tech.integrated <- FindNeighbors(tech.integrated, dims = 1:ndims)
tech.integrated <- FindClusters(tech.integrated, resolution = 0.5)
tech.integrated <- RunUMAP(tech.integrated, dims = 1: ndims)
```
Plot UMAP

```{r warning=FALSE, results='hide'}
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	tech.integrated <- FindClusters(tech.integrated, resolution = x)
}
```

```{r}
for (meta.col in colnames(tech.integrated@meta.data)){
	if(grepl(pattern = "integrated_snn_res", x = meta.col)==TRUE | grepl(pattern = "orig.ident", x = meta.col)==TRUE){
		myplot <- DimPlot(tech.integrated, 
											group.by = meta.col,
											reduction = "umap", 
											pt.size = 1,
											cols = color.palette) + 
			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col) ))
		plot(myplot)
		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap.png"), height = 800, width = 800)
		plot(myplot)
		dev.off()
		myplot <- DimPlot(tech.integrated, 
											group.by = meta.col,
											reduction = "umap", 
											pt.size = 1,
											cols = color.palette) + 
			facet_wrap(meta.col) + 
			ggtitle(paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col)))

		png(filename = paste0(projectName, " pctVar", pct.var, "res.", gsub("integrated_snn_res.", "", meta.col), "-umap_FacetRes.png"), height = 800, width = 800)
		plot(myplot)
		dev.off()

	}
}
```
### Evaluate cluster stability
```{r fig.height=5}
clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") + 
	scale_color_continuous(low = 'red3', high = 'white')
png(filename = paste0(projectName, "_pctVar", pct.var, "-clustree.png"), height = 800, width = 1600)
clustree(tech.integrated, prefix = "integrated_snn_res.", node_colour = "sc3_stability") +
	scale_color_continuous(low = 'red3', high = 'white')
dev.off()
```
```{r}
saveRDS(tech.integrated, file = paste0(projectName, "_pctVar", pct.var, ".RDS"))
```

# Pick set of conditions to investigate further
Now that have checked multiple pct.vars and object.res's, pick the favourite combination and do more detailed characerization of the. This includes:
* Stats on each cluster
* Find markers
* Export lists for GSEA

## Reload object 
```{r}
tech.integrated <- readRDS(paste0(projectName, "_pctVar", fav.var, ".RDS"))
tech.integrated <- SetIdent(tech.integrated, value = paste0("integrated_snn_res", object.res))
```
for each resolution, number/percentage of cells in each cluster?
```{r}
current_res <- 'integrated_snn_res.0.5'
tech.integrated@meta.data$orig.ident[tech.integrated@meta.data$orig.ident == paste0(pop.of.interest, cellranger.folder)] <- "10X"
tech.integrated@meta.data$orig.ident[tech.integrated@meta.data$orig.ident == paste0(pop.of.interest)] <- "biomark"
tech.integrated@meta.data$orig.ident <- factor(tech.integrated@meta.data$orig.ident)

cluster_ids <- sort(unique(tech.integrated@meta.data[,current_res]))
counts_df <- data.frame(matrix(nrow = length(cluster_ids), ncol = 4))
rownames(counts_df) <- cluster_ids
colnames(counts_df) <- c("10X", "biomark", "pct10X", "pctbiomark")
tot.10X <- nrow(tech.integrated@meta.data[tech.integrated@meta.data$orig.ident == "10X", ])
tot.biomark <- nrow(tech.integrated@meta.data[tech.integrated@meta.data$orig.ident == "biomark", ])
for(id in cluster_ids){
	cell_value <- nrow(tech.integrated@meta.data[(tech.integrated@meta.data[current_res] == id) & 
																							 	(tech.integrated@meta.data$orig.ident == "10X"),])
	counts_df[id, "10X"] = cell_value
	counts_df[id, "pct10X"] = round(cell_value/tot.10X*100, 2)
	cell_value <- nrow(tech.integrated@meta.data[(tech.integrated@meta.data[current_res] == id) & 
																							 	(tech.integrated@meta.data$orig.ident == "biomark"),])
	counts_df[id, "biomark"] = cell_value
	counts_df[id, "pctbiomark"] = round(cell_value/tot.biomark*100, 2)
}
```
```{r}
try(xlsx::write.xlsx(x = counts_df, 
								 file = paste0(projectName, " pctVar", pct.var, "_stats.xlsx"), 
								 sheetName = "cluster_stats", 
								 col.names = TRUE, 
								 row.names = TRUE, 
								 append = TRUE)
)
counts_df
```

## Create `FindAllMarkers()` lists for GSEA
```{r}
ident.list <- c(paste0("integrated_snn_res", object.res))

for(tested.ident in ident.list){
	Idents(tech.integrated) <- tested.ident
	all.markers <- FindAllMarkers(tech.integrated)
	try(xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
									 file = paste0(projectName, " pctVar", pct.var, "_allres.xlsx"), 
									 sheetName = tested.ident, 
									 col.names = TRUE, 
									 row.names = FALSE, 
									 append = TRUE)
			)
}
```

## Map HGNC symbols
```{r}
tech.integrated <- SetIdent(tech.integrated, value = paste0("integrated_snn_res", object.res))
object.res.allmarkers <- FindAllMarkers(tech.integrated)
Mouse2HumanTable <- Mouse2Human(object.res.allmarkers$gene)

HGNC <- with(Mouse2HumanTable, Mouse2HumanTable$HGNC[match(object.res.allmarkers$gene, Mouse2HumanTable$MGI)])
head(object.res.allmarkers)
object.res.allmarkers$HGNC <- HGNC
# tail(object.res.allmarkers)
sig.res <- object.res.allmarkers[object.res.allmarkers$p_val_adj <= 0.05, ]
sig.res <- sig.res[c("avg_log2FC", "HGNC", "cluster")]
sig.res <- sig.res[!(sig.res$HGNC == "" | is.na(sig.res$HGNC)),] # GSEA will fail if there are any blanks or NAs in the table
sig.res <- sig.res[]

for(cluster in unique(sig.res$cluster)){
	print(paste("writing cluster", cluster))
	new.table <- sig.res[sig.res$cluster == cluster, c("HGNC", "avg_log2FC")]
	new.table <- new.table[order(-new.table$avg_log2FC), ]
	try(dir.create(paste0("RankList_findAll_hgnc/"), showWarnings = FALSE))
	try(write.table(new.table, file = paste0("RankList_findAll_hgnc/", projectName, "pctVar", pct.var, "res", object.res, "cluster", cluster, ".rnk"), quote = FALSE, row.names = FALSE, col.names = TRUE, sep = "\t", ))

}
```

# Merge index sorting data

## load object
```{r}
tech.integrated <- readRDS(paste0(projectName, "_pctVar", fav.var, ".RDS"))
tech.integrated <- SetIdent(tech.integrated, value = "integrated_snn_res.0.5")
tech.integrated@meta.data$orig.ident[tech.integrated@meta.data$orig.ident == paste0(pop.of.interest, cellranger.folder)] <- "10X"
tech.integrated@meta.data$orig.ident[tech.integrated@meta.data$orig.ident == paste0(pop.of.interest)] <- "biomark"
tech.integrated@meta.data$orig.ident <- factor(tech.integrated@meta.data$orig.ident)
```

## load index sort data
```{r}
mfi <- read.csv(file = paste0(pop.of.interest, "_mfi_unNormalized.txt"), sep = "\t", row.names = 1, skipNul = TRUE)
head(mfi)
```

sort_dict = {'010319':['p1', 'p3'], '121218': ['p2', 'p4'], '020119': ['p6', 'p7'], '090618': ['p12', 'p13']}
normalize based on sort data  
Plot to check distribution
```{r}
for(colname in colnames(mfi[,grepl("pos$", names(mfi))])){
	h <- hist(mfi[, colname], plot = FALSE) 
	h$density = h$counts/sum(h$counts)*100
	plot(h,freq=FALSE, main = colname, xlab = "mfi", ylab = "% Cells")
	}
```
So some distributions are gaussian, and some are poisson. Best bet is probaby just to normalize everyone.  
> Min-max normalization: guarantees the same scale, but is thrown by outliers  
> Z score normalization: does not guarantee the same scale, but handles outliers better

Flow is full of outliers. Going to z-score normalize

```{r}
fluor.cols <- names(mfi[, grepl("pos$", names(mfi))])
mfi.norm <- data.frame(matrix(nrow = 0, ncol = length(fluor.cols)))
colnames(mfi.norm) <- fluor.cols
for(sortdate in unique(mfi$sortdate)){
	mfi.subset <- scale(mfi[mfi$sortdate == sortdate, fluor.cols])
	mfi.norm <- rbind(mfi.norm, mfi.subset)
}
mfi.norm["pid"] <- do.call(rbind, strsplit(rownames(mfi.norm), '_'))[,3]
```
Add metadata column
```{r}
tech.integrated <- AddMetaData(tech.integrated, metadata = mfi.norm)
```
now plot some stuff
```{r}
biomark.integrated <- subset(tech.integrated, subset = orig.ident == "biomark")
```
```{r fig.height=12, fig.width=5}
RidgePlot(biomark.integrated, features = fluor.cols, cols = color.palette, ncol = 1)
```
```{r}
png(filename = paste0(projectName, " pctVar", fav.var, "res.", object.res, "-fluor.png"), height = 2400, width = 800)
RidgePlot(biomark.integrated, features = fluor.cols, cols = color.palette, ncol = 1)
dev.off()
```