### 2021-11-08 MergeBiomarkWith10X/MergeBiomark281With10X.Rmd

---
title: "MergeBiomark281With10X"
output: html_notebook
---

```{r setup}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/MergeBiomarkWith10X/")
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
```

# Set global variables
```{r}
projectName <- "MergeBiomark281With10X"
jackstraw.dim <- 40
```

```{r}
source("msAggr_AnalysisCode/read_10XGenomics_data.R")
source("msAggr_AnalysisCode/PercentVariance.R")
source("msAggr_AnalysisCode/Mouse2Human_idconversion.R")
```

store session info
```{r }
sink(paste0(projectName, ".txt"))
sessionInfo()
sink()
```

# Create seurat object

## Load biomark data
Try to read in Biomark Data

Will need to include:
cell list
probe list
expression matrix

```{r}
biomark.data <- read.table("Biomark281_expressionMatrix.txt", sep = "\t", header = TRUE, row.names = 1)
biomark.data[is.na(biomark.data)] <- 0
biomark.data.T <- data.table::transpose(biomark.data)
colnames(biomark.data.T) <- rownames(biomark.data)
rownames(biomark.data.T) <- colnames(biomark.data)
# fix non-MGI gene names
rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd117'] <- 'Kit'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd123'] <- 'Il3ra'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Aa'] <- 'H2-Aa'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Ab1'] <- 'H2-Ab1'
```

Pull in original cellID info

```{r}
biomark.order <- read.table("origBiomark_cellIDorder.txt", sep = "\t", header = FALSE, row.names = 1)
biomark.order["clust.ID"] <- NA
biomark.order[1:49, "clust.ID"] <- 0
biomark.order[50:87, "clust.ID"] <- 17
biomark.order[88:102, "clust.ID"] <- 10
biomark.order[103:117, "clust.ID"] <- 0
biomark.order[118:127, "clust.ID"] <- 3
biomark.order[128:150, "clust.ID"] <- 0
biomark.order[151:189, "clust.ID"] <- 11
biomark.order[190:281, "clust.ID"] <- 0
```

```{r}
fluor.table <- biomark.data[grepl("pos", colnames(biomark.data))]
```

```{r warning=FALSE, message=FALSE}
biomark.meta <- data.frame(stringr::str_split_fixed(rownames(biomark.data), "_", 3)[,2:3], row.names = rownames(biomark.data))
colnames(biomark.meta) <- c("wid", "pid")

biomark.meta['nFeature_RNA'] <- NA
for(cell in colnames(biomark.data.T)){
	ngenes <- length(which(!(biomark.data.T[colnames(biomark.data.T > 0) == cell])))
	biomark.meta[rownames(biomark.meta) == cell, 'nFeature_RNA'] <- ngenes
}
i1 <- match(row.names(biomark.meta), row.names(biomark.order))
biomark.meta[names(biomark.order)[1]] <- lapply(biomark.order[1], `[`, i1)

i1 <- match(row.names(biomark.meta), row.names(fluor.table))
biomark.meta[names(fluor.table)[1:ncol(fluor.table)]] <- lapply(fluor.table[1:ncol(fluor.table)], `[`, i1)

```

```{r}
# N.B., can't find alternate name for Kng2, so remove
biomark.data.T <- biomark.data.T[!(rownames(biomark.data.T) == "Kng2"), ]
biomark.data.T <- as.sparse(biomark.data.T)
biomark <- CreateSeuratObject(biomark.data.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
```

## Load CMP data

```{r warning=FALSE}
setwd("../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
object.data <-Read10X(data_file.list)
cmp.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = "10X")
```

Remove non-biomark genes from cmp.object

```{r}
biomark.genes <- rownames(biomark@assays$RNA@data)
cmp.object <- subset(cmp.object, features = biomark.genes)
cmp.object@meta.data["wid"] <- '10X'
cmp.object@meta.data["pid"] <- "p14"
setdiff(rownames(biomark@assays$RNA@data), rownames(cmp.object@assays$RNA@data))
```

Clean up to free memory

```{r}
remove(object.data, biomark.data, biomark.meta)
```

## merge objects
```{r}
seurat.object <- merge(cmp.object, biomark, project = 'mergebiomark')
seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMPm2'] <- '10X'
seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMP'] <- 'Biomark'
seurat.object
```

```{r}
seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 10)
```

Scale data (linear transformation)

```{r}
all.genes <- rownames(seurat.object)
seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("orig.ident", "nFeature_RNA", "pid", "nCount_RNA"))
```

### Save merged object
```{r}
# saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
```
## Run PCA
```{r}
seurat.object <- RunPCA(seurat.object, ndims.print = 1:5, nfeatures.print = 5, features = all.genes)
```

```{r}
DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)

```

Calculate dimensionality
```{r, figures-side, fig.show='hold', out.width="50%"}
ElbowPlot(seurat.object, ndims = 50)
percent.variance(seurat.object@reductions$pca@stdev)
```

## Calculate percent variance
```{r}
tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))

```

## Add cluster IDs from Seurat v1

Exported cell IDs for clusters 3, 17, 10, 11 from Seurat v1. Will add these IDs as a metadata column.  
Create column "clust.ID" and populate with 0's. Then import IDs for clusters

```{r}
clust3.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust17.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust10.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust11.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
```

## map cluster ids from Seuratv1
```{r}
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust3.cells] <- 3
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust17.cells] <- 17
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust10.cells] <- 10
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust11.cells] <- 11
seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
seurat.object@meta.data$clust.ID <- as.factor(seurat.object@meta.data$clust.ID)
```

## Lets see which cells correspond to our old clusters

```{r}
DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
```

```{r fig.width=15, fig.height=5}
DimPlot(seurat.object, split.by = "clust.ID", group.by = "orig.ident", pt.size = 2)
```

## Split object for anchor-based integration

```{r}
cmp.subset <- SplitObject(seurat.object, split.by = "orig.ident")$`10X`
DimPlot(cmp.subset, group.by = "clust.ID")
```

## Follow integration protocol

https://satijalab.org/seurat/archive/v3.0/integration.html

```{r}
tech.list <- SplitObject(seurat.object, split.by = "orig.ident")

for(i in 1:length(tech.list)){
	tech.list[[i]] <- NormalizeData(tech.list[[i]], verbose = FALSE)
	tech.list[[i]] <- FindVariableFeatures(tech.list[[i]], selection.method = "vst", nfeatures = 10, verbose = FALSE)
}
```

## identify anchors
```{r}
# reference.list <- tech.list["10X"]
tech.anchors <- FindIntegrationAnchors(object.list = tech.list, dims = 1:30, anchor.features = all.genes, k.anchor = 20, reduction = 'rpca', k.filter = NA)
# tech.integrated <- IntegrateData(anchorset = tech.anchors, dims = 1:30, k.weight = 20, normalization.method = "SCT")
tech.integrated <- IntegrateData(anchorset = tech.anchors, dims = 1:30, k.weight = 5)
```
## tech.integrated PCA

```{r}
DefaultAssay(tech.integrated) <- "integrated"
tech.integrated <- ScaleData(tech.integrated)
tech.integrated <- RunPCA(tech.integrated, npcs = 30)
DimPlot(tech.integrated, reduction = "pca", group.by = "orig.ident")
```

so it looks like anchor-based integration is much nicer! goign to stick with this approach.

Problem now is, how do we find whether Biomark overlays on 10X

```{r fig.height = 5, fig.width=15}
DimPlot(tech.integrated, split.by = "clust.ID", shape.by = "orig.ident", pt.size = 2, cols = c("black", "red"), reduction = "pca")
```

```{r fig.height=10, fig.width=10}
png(filename = "tech.integrated_biomark281Fluor.png", height = 1600, width = 1600)
FeaturePlot(tech.integrated, features = c("CD34pos", "cKitpos", "CD9pos", "CD36pos", "CD41pos", "CD48pos", "CD123pos", "CD150pos", "Flk2pos"), split.by = "clust.ID", reduction = "pca", cols = c("gray92", "blue"), keep.scale = "all", ncol = 1)
dev.off()

```

## tech.integrated umap
```{r}
tech.integrated <- RunUMAP(tech.integrated, reduction = "pca", dims = 1:30)
```
## Plot UMAP
```{r fig.width=15, fig.height=5}
DimPlot(tech.integrated, split.by = "clust.ID", shape.by = "orig.ident", pt.size = 2, cols = c("black", "red"), reduction = "umap")
```

# Biomark Only
## Load biomark data

```{r}
biomark.data <- read.table("Biomark281_expressionMatrix.txt", sep = "\t", header = TRUE, row.names = 1)
biomark.data[is.na(biomark.data)] <- 0
biomark.data.T <- data.table::transpose(biomark.data)
colnames(biomark.data.T) <- rownames(biomark.data)
rownames(biomark.data.T) <- colnames(biomark.data)
# fix non-MGI gene names
rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd117'] <- 'Kit'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'Cd123'] <- 'Il3ra'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Aa'] <- 'H2-Aa'
rownames(biomark.data.T)[rownames(biomark.data.T) == 'H2.Ab1'] <- 'H2-Ab1'
```

Pull in original cellID info

```{r}
biomark.order <- read.table("origBiomark_cellIDorder.txt", sep = "\t", header = FALSE, row.names = 1)
```

```{r}
biomark.order["clust.ID"] <- NA
biomark.order[1:49, "clust.ID"] <- 0
biomark.order[50:87, "clust.ID"] <- 17
biomark.order[88:102, "clust.ID"] <- 10
biomark.order[103:117, "clust.ID"] <- 0
biomark.order[118:127, "clust.ID"] <- 3
biomark.order[128:150, "clust.ID"] <- 0
biomark.order[151:189, "clust.ID"] <- 11
biomark.order[190:281, "clust.ID"] <- 0
```

```{r warning=FALSE, message=FALSE}
biomark.meta <- data.frame(stringr::str_split_fixed(rownames(biomark.data), "_", 3)[,2:3], row.names = rownames(biomark.data))
colnames(biomark.meta) <- c("wid", "pid")

biomark.meta['nFeature_RNA'] <- NA
for(cell in colnames(biomark.data.T)){
	ngenes <- length(which(!(biomark.data.T[colnames(biomark.data.T > 0) == cell])))
	biomark.meta[rownames(biomark.meta) == cell, 'nFeature_RNA'] <- ngenes
}
i1 <- match(row.names(biomark.meta), row.names(biomark.order))
biomark.meta[names(biomark.order)[1]] <- lapply(biomark.order[1], `[`, i1)

```

```{r}
# N.B., can't find alternate name for Kng2, so remove
biomark.data.T <- biomark.data.T[!(rownames(biomark.data.T) == "Kng2"), ]
biomark.data.T <- as.sparse(biomark.data.T)
biomark <- CreateSeuratObject(biomark.data.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
```

```{r}
biomark <- NormalizeData(biomark, normalization.method = "LogNormalize", scale.factor = 10000)
biomark <- FindVariableFeatures(biomark, selection.method = "vst", nfeatures = 10)
all.genes <- rownames(biomark)
biomark <- ScaleData(biomark, features = all.genes, vars.to.regress = c("nFeature_RNA", "pid", "nCounts_RNA"))
```

### Save merged object
```{r}
# saveRDS(biomark, file = paste0(projectName, "_raw.RDS"))
```
## Run PCA
```{r}
biomark <- RunPCA(biomark, ndims.print = 1:5, nfeatures.print = 5, features = all.genes)
```

```{r}
DimPlot(biomark, reduction = "pca", group.by = "orig.ident")
VizDimLoadings(biomark, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)

```

Calculate dimensionality
```{r}
ElbowPlot(biomark, ndims = 50)
percent.variance(biomark@reductions$pca@stdev)
```

## Calculate percent variance
```{r}
tot.var <- percent.variance(biomark@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))

```

## Add cluster IDs from Seurat v1

Exported cell IDs for clusters 3, 17, 10, 11 from Seurat v1. Will add these IDs as a metadata column.  
Create column "clust.ID" and populate with 0's. Then import IDs for clusters

## Lets see which cells correspond to our old clusters

```{r}
DimPlot(biomark, reduction = "pca", group.by = "orig.ident") + facet_wrap(~orig.ident)
```

```{r fig.width=15, fig.height=8}
biomark@meta.data$clust.ID <- as.factor(biomark@meta.data$clust.ID)
DimPlot(biomark, group.by = "clust.ID", pt.size = 2, shape.by = "clust.ID")
```

```{r fig.width=15, fig.height = 5}
DimPlot(biomark, split.by = "clust.ID", pt.size = 2, shape.by = "orig.ident")
```