### 2021-12-10 monocle/msAggr_monocle.Rmd
Added msAggr document

---
title: "msAggr_monocle"
output: html_document
---

# Setup
## Load libraries
```{r setup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/Monocle/")
library(monocle3)
library(ggplot2)
library(dplyr)
```
## Source functions
```{r}
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
```
# Load datasets
Provide list of datasets
```{r}
data.list <- c("/Users/heustonef/Desktop/10XGenomicsData/cellRanger/LSKm2",
							"/Users/heustonef/Desktop/10XGenomicsData/cellRanger/CMPm2", 
							"/Users/heustonef/Desktop/10XGenomicsData/cellRanger/MEPm", 
							"/Users/heustonef/Desktop/10XGenomicsData/cellRanger/GMPm")
```