### 2021-12-10 monocle/CMP_monocle.Rmd
Add CMP markdown and import cellranger data as cds

---
title: "CMP_monocle"
output: html_notebook
---
# Setup
## Define notebook variables
```{r}
working.genome <- "mm10"
umi.cutoff <- 200
```

## Load libraries
```{r setup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/Monocle/")
library(monocle3)
library(ggplot2)
library(dplyr)
```
## Source functions
```{r}
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
```
# Load datasets
## Provide list of datasets
```{r}
data.path <- c("/Users/heustonef/Desktop/10XGenomicsData/cellRanger/CMPm2")
cds <- load_cellranger_data(data.path)
# cds <- new_cell_data_set(file.path(data.path, "filtered_feature_bc_matrix", "matrix.mtx.gz"),
# 										cell_metadata = cell_metadata, 
# 										gene_metadata = read.delim(file.path(data.path, "filtered_feature_bc_matrix"), header = FALSE, sep = "\t")[, 1:2]
# 										)
```
Extract metadata and gene_short_names from seurat.object
```{r, eval=FALSE}
# Don't run
library(Seurat)
seurat.object <- readRDS("../msAggr_seurat/msAggr_seurat_dim26.RDS")
cell_metadata <- seurat.object@meta.data

```