### 2021-12-10 monocle/IndividualPops_monocle.Rmd
Changed CMP_monocle to IndividualPops_monocle because duh everyone should be analyzed the same

---
title: "CMP_monocle"
output: html_notebook
---
# Setup
## Define notebook variables
```{r}
working.genome <- "mm10"
umi.cutoff <- 200
data.path <- "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/IndividualPops/"
cell.pop <- "CMP"
```

## Load libraries
```{r setup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/Monocle/")
library(monocle3)
library(ggplot2)
library(dplyr)
```
## Source functions
```{r}
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
```
# Load datasets
## Provide list of datasets
After messing with this a bit, I've decided the easiest thing to do is integrate the data from the seurat object. This object already has the metadata and gene_short_names that I will need, and based on monocle3 documentation may be the easiest way to make sure everything gets where it needs to be.
```{r, eval=FALSE}
# Don't run
library(Seurat)
data.path <- 
seurat.object <- readRDS("../msAggr_seurat/msAggr_seurat_dim26.RDS")

cell_metadata <- seurat.object@meta.data

```